---
output:
  github_document:
    pandoc_args: --webtex
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# DRI-GLUCoSE

<!-- badges: start -->
<!-- badges: end -->

As elaborated in our recent analyses (Walker et al. 2019; Scarpone et al. 2020), nearly all previous studies in the literature use either census unit boundaries or simple buffer zones to measure an individual’s built environment (BE) exposures or to characterize their local socioeconomic status (SES) (Rhew et al. 2011, Gong et al. 2014, Fuertes et al. 2014). Therefore, we present a distance-weighted, network-based model for quantifying the combined effects of local greenspace and SES on diabetes risk, from which we derive an area-based Diabetes Risk Index of Greenspace, Land Use and Socioeconomic Environments (DRI-GLUCoSE).\
The goal of the DRIGLUCoSE package is to provide a public package containing functions and code used in the development of the DRI-GLUCoSE Index.

## Installation

You can install the latest version of DRIGLUCoSE from [GitHub](https://CRAN.R-project.org) with:

```{r eval=FALSE, include=TRUE}
devtools::install_git("https://github.com/STBrinkmann/DRIGLUCoSE")
```


## Usage

Once installed, the library can be loaded as follows:

```{r message=FALSE, warning=FALSE}
library(DRIGLUCoSE)
```


# Methods
One key purpose of this package is, to provide functions for route networked derived isochrones. For that purpose we have provided a simple sf object of 2 points in Erlangen, Germany.

```{r}
data(Erlangen)
Erlangen
```

## Census variables
In our analysis we acquired data of the Canadian census dissemination areas (DA, the smallest available census area with an average population of 400-800 residents) for the census year 2006 from the federal government agency Statistics Canada. It has been converted to a shapefile (sf) with one column per census variable. To demonstrate we use the following randomly generated data:

```{r}
set.seed(1234)
census <- sf::st_make_grid(
  # Use Sample Data and apply 25 minutes buffer (Speed[m/min] * 25[min]) 
  Erlangen %>% dplyr::mutate(geom = sf::st_buffer(geom, Speed*25)),
  cellsize = 100
  ) %>% 
  st_as_sf() %>% 
  mutate(census_var = sample(1000:10000, n(), replace = TRUE)) %>% 
  rename(geom = x)

census
```

## Greenspace
In our analysis we acquired LANDSAT images through the United States Geological Survey’s EarthExplorer platform ([https://earthexplorer.usgs.gov/]). The Normalized Difference Vegetation Index (NDVI) is used as a metric to model greenspace exposure. Pre-processing of the LANDSAT images and NDVI calculation has been conducted using the _LS_L1C_ function:

```{r message=FALSE, warning=FALSE, eval=FALSE, include=TRUE}
DRIGLUCoSE::LS_L1C(l1c_path = "docs/LC08_L1TP_193026_20200423_20200508_01_T1_small/", 
                   out_dir = "docs/LS_PreProcessed",
                   # Use Sample Data and apply 25 minutes buffer (Speed[m/min] * 25[min]) 
                   sf_mask = DRIGLUCoSE::Erlangen %>% 
                     dplyr::mutate(geom = sf::st_buffer(geom, Speed*25)),
                   cores = 20)
```

## Exposure Model
In order to estimate each participant’s potential exposures to greenspace and local SES, we (i) mapped age- and sex-specific walkable zones around their residential address, and (ii) applied a negative logit weighting function, such that the estimated effect of greenspace or SES decreases as distance from the home increases.

### (i) Road network data and isochrones
In order to compute network-based distance metrics, we acquired street data from OpenStreetMap using the R-package osmdata (Padgham et al. 2017). Road types not suitable for walking were removed (e.g., motorways). Network data were topologically corrected and split into ~30 metre-long segments using the R package _nngeo_ (Dorman 2020).

```{r message=FALSE, warning=FALSE, eval=FALSE, include=TRUE}
erlangen.osm <- DRIGLUCoSE::osm_roads(x = Erlangen, dist = 20, 
                                      speed = "Speed", cores = 2)
erlangen.osm
```


This network data was used to derive walking distance buffers for each participant, based on age- and sex-specific walking speed estimates derived from the literature (Dewulf 2012). Starting from each participant’s place of residence, we computed network-constrained buffers with an off-road width of 40 meters, running in 2-minute increments from 0 to 20 minutes, using the A*-algorithm (Hart, Nilsson & Raphael 1968). This therefore resulted in each participant having ten concentric isochrones, the sizes of which are a function of participant age and sex.\
Since the road network contains a lot of features (n=14292), this will take some time (~15-30 minutes).

```{r warning=FALSE, eval=FALSE, include=TRUE}
erlangen.isodistances <- DRIGLUCoSE::isodistances(x = Erlangen, 
                                                  road_network = erlangen.osm, 
                                                  tag = "id", speed = "Speed",
                                                  isochrones_seq = seq(2, 20, 2),
                                                  cores = 2)

erlangen.isochrones <- DRIGLUCoSE::isochrones(x = erlangen.isodistances, 
                                              buffer = 40, cores = 2)
```

```{r fig.height=8, fig.width=11, eval=FALSE, include=TRUE}
erlangen.isodistances %>%
  dplyr::mutate(time = forcats::fct_rev(time)) %>% 
  dplyr::filter(id == 2) %>% 
  ggplot2::ggplot(aes(colour = time)) + 
  ggplot2::geom_sf(size = 1) + 
  ggplot2::scale_color_discrete(name = "Walking-Time\n(minutes)") +
  ggthemes::theme_map() + 
  ggplot2::theme(legend.position = c(0.96, 0.35),
                 legend.title=element_text(size=18), 
                 legend.text=element_text(size=14))
```

## Distance-weighting
In order to account for the diminishing effect of SES and greenspace exposure as distance increases, we fitted a logit function to weight each incremental isochrone, such that the influence of a variable decreases with increasing distance from the household, i.e., features that are farther away have less influence than nearby features, as illustrated in Figure 2. A logit function was selected as it heuristically approximates a suitable distance-decay function (Bauer and Groneberg 2016; Jia et al. 2019) and various parameterisations (for parameters _b_ and _m_) of the logit function were assessed using sensitivity analysis.\
The distance-weighting is separated in two parts, first the logit function (1) that is used for both SES and greenspace variables, and second the proportional weights function (4) that is only applied on SES variables.

$$
\begin{align}
  G_t =
    \begin{cases}
      \cfrac{\int_0^{r_t} g(r)dr}{\int_0^{t_max} g(r)dr}, t=1\\
      \cfrac{\int_{t-1}^{r_t} g(r)dr}{\int_0^{t_max} g(r)dr}, t>1
    \end{cases}
\end{align}
$$
