---
output: github_document
always_allow_html: yes
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# DRI-GLUCoSE

<!-- badges: start -->
<!-- badges: end -->

The goal of DRIGLUCoSE is to provide a public package containing functions and code used in the development of the DRI-GLUCoSE Index.

## Installation

You can install the latest version of DRI.GLUCoSE from [GitHub](https://CRAN.R-project.org) with:

```{r eval=FALSE, include=TRUE}
devtools::install_git("https://github.com/STBrinkmann/DRIGLUCoSE")
```


## Usage

Once installed, the library can be loaded as follows:

```{r}
library(DRIGLUCoSE)
```

## Isochrone workflow

One key purpose of this package is, to provide functions for route networked derived isochrones. For that purpose we have provided a simple sf object of 2 points in Erlangen, Germany.

```{r}
data(Erlangen)
Erlangen
```

The first step is to download and preprocess OSM road data:
```{r eval=FALSE, include=TRUE}
erlangen.osm <- osm_roads(x = Erlangen, dist = 20, speed = "Speed", cores = 2)
```

Next, the isodistances will be calculated based on the road network. Since the road network contains a lot of features (n=14061), this will take some time (~15-30 minutes).
```{r eval=FALSE, include=TRUE}
erlangen.isodistances <- isodistances(x = Erlangen, road_network = erlangen.osm, tag = "id",
                                      isochrones_seq = seq(2, 20, 2), speed = "Speed", cores = 2)
```

Finally, the isochrones can be calculated by applying an off-road buffer on the isodistances. The results can be mapped using the [tmap](https://github.com/mtennekes/tmap) package, however at this point I do not know how to include the .html map into the GitHub-ReadMe. 
```{r eval=FALSE, include=TRUE}
erlangen.isochrones <- isochrones(erlangen.isodistances, 40, 4)
```

```{r eval=FALSE, include=TRUE}
tmap::tmap_mode("view")

tmap1 <- tmap::tm_shape(Erlangen[Erlangen$id == 1,] %>% sf::st_buffer(20*Erlangen[Erlangen$id == 1,]$Speed),
               name = "Euclidean Buffer") +
  tmap::tm_polygons(alpha = 0.1, legend.show = F) +
  tmap::tm_shape(erlangen.isochrones[erlangen.isochrones$tag == 1, ] %>% dplyr::mutate(time = forcats::fct_rev(time))) +
  tmap::tm_polygons(col="time",
                    palette = "Dark2",
                    alpha = 0.1,
                    legend.show = FALSE) +
  tmap::tm_facets(by = "time", as.layers = TRUE, free.coords = FALSE, drop.units = TRUE) +
  tmap::tm_shape(erlangen.isodistances[erlangen.isodistances$tag == 1, ] %>% dplyr::mutate(time = forcats::fct_rev(time))) +
  tmap::tm_lines(col="time",
                 palette = "Dark2",
                 lwd = 3,
                 legend.col.show = FALSE,
                 legend.lwd.show = FALSE) +
  tmap::tm_facets(by = "time", as.layers = TRUE, free.coords = FALSE, drop.units = TRUE) +
  tmap::tm_shape(Erlangen[Erlangen$id == 1,], name = "Centre") +
  tmap::tm_dots(col = "black", size = 0.1) +
  tmap::tm_layout(legend.outside = TRUE) +
  tmap::tm_basemap(leaflet::providers$OpenStreetMap)

tmap1
```













